---
title: "mvse_pipeline"
author: "Toby Koch"
date: '2022-08-11'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, prompt = F, include = F, warning = F)
rm(list = ls())
#create data directories because they are not ignored in the github repository
if(!dir.exists("data_mid")){
  dir.create("data_mid")
}

if(!dir.exists("data_output")){
  dir.create("data_output")
}
```

WHAT SCRIPT BELOW DOES
Get FIPS data for matching with counties you want to keep in the shape file for calculating centroids
```{bash engine.opts ='-l'}
wget -P data_input/ -O data_input/us_fips.csv https://gist.githubusercontent.com/dantonnoriega/bf1acd2290e15b91e6710b6fd3be0a53/raw/11d15233327c8080c9646c7e1f23052659db251d/us-state-ansi-fips.csv
```

BEFORE RUNNING ENSURE THAT THE FOLLOWING FILES ARE IN YOUR data_input folder:
1. countries.csv: COUNTRIES/REGIONS YOU WISH TO FILTER FOR
2. MVSE.R has been downloaded from https://sourceforge.net/projects/mvse/

BEFORE RUNNING ENSURE DATA HAVE BEEN DOWNLOADED IN THE DATA INPUT FOLDER

WHAT SCRIPT BELOW DOES
scripts/weather_data_mvse_prep.R
1. identifies counties with and without Hx of cases
2. identifies counties and there centroids for averaging the weather
3. was using to create lat and long data for the visual crossing data


```{bash}
#pulls in country data to filter stations in your desired locations

scripts/weather_data_mvse_prep.R -m centroid

#note: it sources the function to pull the state but may be 
```

NEXT STEP IS TO USE THE WEATHER_STATION_BATCH.CSV TO UPLOAD TO https://visualcrossing.com 
and put resulting file back into data_input called weather_data.csv

WHAT SCRIPT DOES
scripts/weather_data_mvse_prep.R does the following:
1. reads in climate data from visualcrossing
2. splits by grouping variables determined by the geographical grouping variable (--geo)
3. calculates average temperature, humidity and rainfall
4. Removes negative temperatures
5. filters by desired date range
6. calculate p from month lower to month higher for each year
5. writes files to data_mid to then be run by MVSE_loop.R

#it is in bash because that is how the options parser wants to run
```{bash}
scripts/weather_data_mvse_prep.R --geo centroid --datelower 2000-01-01 --dateupper 2020-12-31 --monthlower 4 --monthupper 10
```

WHAT SCRIPT DOES
1. loads in functions from MVSE package
2. defines paramaters
3. calculates index P
4. writes files to data_mid

```{bash}
scripts/MVSE_loop.R --directory centroid/
```


WHAT SCRIPT BELOW DOES
1. compiles index P output for the various geographic resultions and outputs into one csv file

NOTE that --geo state_year -datelower 2015-01-01 -dateupper 2020-12-31 is the default
```{bash}
scripts/mvse_output.R --geo centroid --datelower 2000-01-01 --dateupper 2020-12-31
```

WHAT SCRIPT BELOW DOES
scripts/cases.R does the following
1. reads in case data for humans and non-humans and combines them

```{r}
source("scripts/cases.R")
```

WHAT SCRIPT BELOW DOES

```{r}
source("scripts/mosquitoes.R")
```

WHAT SCRIPT BELOW DOES
- combines the mvse, mosquito, and case data into one file and creates files grouped by different year variables
- creates monthly data file called 
```{r}
source("scripts/data_combine.R")
```


WHAT SCRIPT DOES
```{r}
source("scripts/data_clean_for_model.R")
```

WHAT SCRIPT BELOW DOES

```{r}
source("scripts/modeling.R")
```

